Optimizing and Incrementalizing Collection Queries by AST Transformations

### 2017-02-02 pldi14/sec-rw.tex
Move PLDI14 RW as a chapter in the second part

% Emacs, this is -*- latex -*-!
\section{Related work}
\label{sec:rw}

@@

\subsection{Dynamic approaches}
One of the most advanced dynamic approach to incrementalization is
self-adjusting computation, which has been applied to Standard ML
and large subsets of C~\citep{Acar09,Hammer11}.
In this approach, programs execute on the original
input in an enhanced runtime environment that tracks the
dependencies between values in a \emph{dynamic
  dependence graph}~\citep{Acar06}; intermediate results are
memoized.
Later, changes to the input propagate through
dependency graphs from changed inputs to results,
updating both intermediate and final results;
this processing is often more efficient than recomputation.

@@

\subsection{Static approaches}
\pg{If we discuss partial evaluation, we should compare to
  \citep{Sundaresh91}}
Static approaches analyze a program at compile-time and produce an
incremental version that efficiently updates the output
of the original program according to changing inputs.

@@

\subsubsection{Finite differencing}
\label{sec:finite-diff}
\citet{Paige82FDC} present derivatives for a first-order language
with a fixed set of primitives.
\citet{Blakeley:1986:EUM} apply these ideas to a class of relational queries.
The database community extended
this work to queries on relational data, such as in \emph{algebraic
  differencing}~\citep{Gupta99MMV}, which inspired our work and
terminology. However, most of this work does not apply to nested
collections or algebraic data types, but only to relational
(flat) data, and no previous approach handles first-class
functions. Incremental support is typically designed
monolithically for a whole language, rather than piecewise.
Improving on algebraic differencing, \citet{Koch10IQE}
\emph{guarantees} asymptotic speedups with a compositional query
transformation and delivers huge speedup in realistic benchmarks,
though still for a first-order database language.

@@

\subsubsection{Static memoization}
\label{ssec:staticmemo}
\citeauthor{Liu00}'s work~\citep{Liu00} allows to incrementalize a first-order base
program $f(\Old{x})$ to compute $f(\New{x})$, knowing how
$\New{x}$ is related to $\Old{x}$. To this end, they transform
$f(\New{x})$ into an incremental program which reuses the
intermediate results produced while computing $f(\Old{x})$, the
base program. To this end, (i) first the base program is
transformed to save all its intermediate results, then (ii) the
incremental program is transformed to reuse those intermediate
results, and finally (iii) intermediate results which are not
needed are pruned from the base program. However, to reuse
intermediate results, the incremental program must often be
rearranged, using some form of equational reasoning, into some
equivalent program where partial results appear literally. For
instance, if the base program $f$ uses a left fold to sum the
elements of a list of integers $\Old{x}$, accessing them from the
head onwards, and $\New{x}$ prepends a new element $h$ to the
list, at no point does $f(\New{x})$ recompute the same results.
But since addition is commutative on integers, we can rewrite
$f(\New{x})$ as $f(\Old{x}) + h$. The author's CACHET system will
try to perform such rewritings automatically, but it is not
guaranteed to succeed. Similarly, CACHET will try to synthesize
any additional results which can be computed cheaply by the base
program to help make the incremental program more efficient.

>>>>

% Emacs, this is -*- latex -*-!
\chapter{Related work}
\label{sec:rw}

@@

\section{Dynamic approaches}
One of the most advanced dynamic approach to incrementalization is
self-adjusting computation, which has been applied to Standard ML
and large subsets of C~\citep{Acar09,Hammer11}.
In this approach, programs execute on the original
input in an enhanced runtime environment that tracks the
dependencies between values in a \emph{dynamic
  dependence graph}~\citep{Acar06}; intermediate results are
memoized.
Later, changes to the input propagate through
dependency graphs from changed inputs to results,
updating both intermediate and final results;
this processing is often more efficient than recomputation.

@@

\section{Static approaches}
\pg{If we discuss partial evaluation, we should compare to
  \citep{Sundaresh91}}
Static approaches analyze a program at compile-time and produce an
incremental version that efficiently updates the output
of the original program according to changing inputs.

@@

\subsection{Finite differencing}
\label{sec:finite-diff}
\citet{Paige82FDC} present derivatives for a first-order language
with a fixed set of primitives.
\citet{Blakeley:1986:EUM} apply these ideas to a class of relational queries.
The database community extended
this work to queries on relational data, such as in \emph{algebraic
  differencing}~\citep{Gupta99MMV}, which inspired our work and
terminology. However, most of this work does not apply to nested
collections or algebraic data types, but only to relational
(flat) data, and no previous approach handles first-class
functions. Incremental support is typically designed
monolithically for a whole language, rather than piecewise.
Improving on algebraic differencing, \citet{Koch10IQE}
\emph{guarantees} asymptotic speedups with a compositional query
transformation and delivers huge speedup in realistic benchmarks,
though still for a first-order database language.

@@

\subsection{Static memoization}
\label{ssec:staticmemo}
\citeauthor{Liu00}'s work~\citep{Liu00} allows to incrementalize a first-order base
program $f(\Old{x})$ to compute $f(\New{x})$, knowing how
$\New{x}$ is related to $\Old{x}$. To this end, they transform
$f(\New{x})$ into an incremental program which reuses the
intermediate results produced while computing $f(\Old{x})$, the
base program. To this end, (i) first the base program is
transformed to save all its intermediate results, then (ii) the
incremental program is transformed to reuse those intermediate
results, and finally (iii) intermediate results which are not
needed are pruned from the base program. However, to reuse
intermediate results, the incremental program must often be
rearranged, using some form of equational reasoning, into some
equivalent program where partial results appear literally. For
instance, if the base program $f$ uses a left fold to sum the
elements of a list of integers $\Old{x}$, accessing them from the
head onwards, and $\New{x}$ prepends a new element $h$ to the
list, at no point does $f(\New{x})$ recompute the same results.
But since addition is commutative on integers, we can rewrite
$f(\New{x})$ as $f(\Old{x}) + h$. The author's CACHET system will
try to perform such rewritings automatically, but it is not
guaranteed to succeed. Similarly, CACHET will try to synthesize
any additional results which can be computed cheaply by the base
program to help make the incremental program more efficient.
